BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications

### 2017-06-28 Paper2/BOAST.tex
Updated Bibliography to use Harvard style.

\documentclass{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{xspace}

@@

\bibliographystyle{IEEEtran}

@@

\markboth{International Journal of Scientific Computing}
{Videau \MakeLowercase{\textit{et al.}}: BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
%\address[mymainaddress]{LIG/CNRS}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code.}, 
        label={lst:BOAST_algebra},
        numbers=left }
\begin{lstlisting}
i = BOAST::Int("i") # or BOAST::Variable::new("i", Int)
k = BOAST::Int("k", :size => 8)
l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 21) ], :local => true )
BOAST::decl i, k, l
BOAST::pr i === 5
j = i + 5
BOAST::pr k === j * 2
BOAST::pr l[k] === 1.0
BOAST::register_funccall("sin")
BOAST::pr l[k+1] === BOAST::sin(j)
\end{lstlisting}

@@

In the Kohn-Sham (KS) formulation of DFT, the electrons are associated to
wavefunctions (orbitals), which are represented by arrays of floating point
numbers. In wavelets formalism, the operators are written via convolutions with
short, separable filters.  The detailed description of how these operations are
defined is beyond the scope of this paper and can be found in the BigDFT
reference paper~\cite{genovese2008daubechies}. Convolutions are basic operations
of lots of scientific application codes, like for example finite differences
approaches, which are universally used in computational physics.

@@

The interest in having a robust and optimally tuned library in this scientific
field is therefore evident. Techniques are under investigation to also provide
end users with fine-tuned binaries rather than the source codes, such that more
aggressive inter procedural optimization can be performed. Indeed, BOAST finds
the optimal source code for a given kernel and compiler configuration but one
could imagine using binary optimizer to the compiled binaries. Those optimizers
could be coupled to BOAST and the output of the process would be the final
binary rather than the source code. This experience therefore would be a first
step toward the release of a tunable optimized convolution library oriented to
computational physics communities.

@@

The most convenient way to obtain an application that can be auto-tuned on a
given platform is to base this application on a widely used computing library.
BLAS~\cite{dongarra1990set} and LAPACK~\cite{laug} are such libraries. These
libraries are either hand tuned for selected platforms or have auto-tuned
implementations. Atlas~\cite{whaley04} is an auto-tuned implementation of
BLAS/LAPACK. ATLAS authors defined the \emph{Automated Empirical Optimization of
Software} methodology that we implemented with BOAST. Their kernel generation is
performed using macro-functions in C.

>>>>

\documentclass{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\usepackage{xspace}
\usepackage{natbib}

@@

\bibliographystyle{SageH}

@@

\markboth{The International Journal of High Performance Computing Applications}
{Videau \MakeLowercase{\textit{et al.}}: BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
%\address[mymainaddress]{LIG/CNRS}

@@

\begin{figure}
\lstset{style=BOAST,
        caption={BOAST code.}, 
        label={lst:BOAST_algebra},
        numbers=right }
\begin{lstlisting}
i = BOAST::Int("i") # or BOAST::Variable::new("i", Int)
k = BOAST::Int("k", :size => 8)
l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 21) ], :local => true )
BOAST::decl i, k, l
BOAST::pr i === 5
j = i + 5
BOAST::pr k === j * 2
BOAST::pr l[k] === 1.0
BOAST::register_funccall("sin")
BOAST::pr l[k+1] === BOAST::sin(j)
\end{lstlisting}

@@

In the Kohn-Sham (KS) formulation of DFT, the electrons are associated to
wavefunctions (orbitals), which are represented by arrays of floating point
numbers. In wavelets formalism, the operators are written via convolutions with
short, separable filters.  The detailed description of how these operations are
defined can be found in the BigDFT
reference paper~\cite{genovese2008daubechies}. Convolutions are basic operations
of lots of scientific application codes, like for example finite differences
approaches, which are universally used in computational physics.

@@

The interest in having a robust and optimally tuned library in this scientific
field is therefore evident. Techniques are under investigation to also provide
end users with fine-tuned binaries rather than the source codes, such that more
aggressive inter procedural optimization can be performed. Indeed, BOAST finds
the optimal source code for a given kernel and compiler configuration but one
could imagine using binary optimizer to the compiled binaries. Those optimizers
could be coupled to BOAST and the output of the process would be the final
binary rather than the source code. This experience therefore would be a first
step toward the release of a tunable optimized convolution library oriented toward
computational physics.

@@

The most convenient way to obtain an application that can be auto-tuned on a
given platform is to base this application on a widely used computing library.
BLAS~\cite{dongarra1990set} and LAPACK~\cite{laug} are such libraries. These
libraries are either hand tuned for selected platforms or have auto-tuned
implementations. Atlas~\cite{whaley04} is an auto-tuned implementation of
BLAS/LAPACK. ATLAS authors defined the \emph{Automated Empirical Optimization of
Software} methodology that we implemented with BOAST. Their kernel generation is
performed using C macro-functions.
